---
title: "School test score analysis"
author: "MaryJo Webster"
date:  "Last updated: `r Sys.Date()`"
output:
  html_document: 
    toc: true
    to_depth: 1
    toc_float: true
    theme: flatly
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning=FALSE, message=FALSE, cache=FALSE, fig.height = 6, fig.width = 8)




#http://strib-data-internal.s3-us-west-1.amazonaws.com/projects/schools/testscore_analysis.html

library(DT)
```


```{r, code=readLines("process_test_score_data.R"), echo=FALSE, warning=FALSE, message=FALSE}

```

```{r}
#these replace column names that get created when using spread() function in several spots below
#update each year
xthisyear <-  'x18_19'
xlastyear <-  'x17_18'
```



```{r}

#overall proficiency rates in reading statewide (for graphic)
state_read <- statewide %>%  filter(data_year>=firstyear, subject=='R',report_description=='All Students') %>% group_by(data_year, subject) %>% 
  summarise(tottested=sum(count_tested), totlevel1=sum(count_level1), totlevel2=sum(count_level2), totlevel3=sum(count_level3), totlevel4=sum(count_level4)) %>% mutate(countproficient=totlevel3+totlevel4, pctproficient=countproficient/tottested)
  
#overall proficiency rates in math statewide (for graphic)
state_math <- statewide %>%  filter(data_year>=firstyear,subject=='M',report_description=='All Students') %>% group_by(data_year, subject) %>% 
  summarise(tottested=sum(count_tested), totlevel1=sum(count_level1), totlevel2=sum(count_level2),  totlevel3=sum(count_level3), totlevel4=sum(count_level4)) %>% mutate(countproficient=totlevel3+totlevel4, pctproficient=countproficient/tottested)

#write.csv(state_read, "./output/state_reading_proficiency.csv", row.names = FALSE)
#write.csv(state_math, "./output/state_math_proficiency.csv", row.names=FALSE)


```

# Statewide
##reading
```{r}
state_read %>% mutate(pctprof=round(pctproficient*100,1), pctlev1=round((totlevel1/tottested)*100,1), pctlev2=round((totlevel2/tottested)*100,1), pctlev3=round((totlevel3/tottested)*100,1), pctlev4=round((totlevel4/tottested)*100,1)) %>% select(data_year, subject, pctprof, pctlev1, pctlev2, pctlev3, pctlev4)  %>%  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```

## math
```{r}
state_math %>% mutate(pctprof=round(pctproficient*100,1), pctlev1=round((totlevel1/tottested)*100,1), pctlev2=round((totlevel2/tottested)*100,1), pctlev3=round((totlevel3/tottested)*100,1), pctlev4=round((totlevel4/tottested)*100,1)) %>% select(data_year, subject, pctprof, pctlev1, pctlev2,  pctlev3, pctlev4)%>%  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```

## Race groups - math
```{r}

statewide <-  statewide %>% mutate(report_category==str_trim(report_category))

state_race <-  statewide %>% filter(data_year>=race_groups_start_year, report_category=='Race/Ethnicity' | report_category=='All Categories', subject=='M') %>%  group_by(subject, data_year, report_description) %>% summarise(tot=sum(count_tested), prof = sum(count_level3)+sum(count_level4)) %>% mutate(pct= (prof/tot)) %>% select(data_year, report_description, pct) %>% filter(report_description!='P-Native Hawaiian/Pacific Islander')

g_race_math <- ggplot(data=state_race, aes(x = state_race$data_year, y = pct, group=report_description, col=report_description)) +
                  geom_line(stat="identity",  size=1)+
  scale_y_continuous(name="Percent proficient", limits=c(0, .8), labels=percent)+
  theme(legend.key.size = unit(.3, "cm"), legend.position="top", legend.title = element_blank())+
  scale_x_discrete(name="School year")+
    theme_hc()+
      labs(title = "Achievement in math", 
       subtitle = "",
       caption = "Graphic by MaryJo Webster")
plot(g_race_math)
```

## Race groups- reading
```{r}
state_race_read <-  statewide %>% filter(data_year>=race_groups_start_year,report_category=='Race/Ethnicity' | report_category=='All Categories', subject=='R') %>%  group_by(subject, data_year, report_description) %>% summarise(tot=sum(count_tested), prof = sum(count_level3)+sum(count_level4)) %>% mutate(pct= (prof/tot)) %>% select(data_year, report_description, pct) %>% filter(report_description!='P-Native Hawaiian/Pacific Islander')

g_race_read <- ggplot(data=state_race_read, aes(x = state_race_read$data_year, y = pct, group=report_description, col=report_description)) +
                  geom_line(stat="identity",  size=1)+
  scale_y_continuous(name="Percent proficient", limits=c(0, .9), labels=percent)+
  scale_x_discrete(name="School year")+
  theme(legend.key.size = unit(.3, "cm"), legend.position="top", legend.title = element_blank())+
    theme_hc()+
      labs(title = "Achievement in reading", 
       subtitle = "",
       caption = "Graphic by MaryJo Webster")
plot(g_race_read)
```


```{r}
#export the data from the above two graphics for our graphics desk
#first widen them out
state_race_read_graphic <-  spread(state_race_read, report_description, -subject)
state_race_math_graphic <-  spread(state_race, report_description, -subject)

write.csv(state_race_read_graphic, './output/state_race_read.csv', row.names=FALSE)
write.csv(state_race_math_graphic, './output/state_race_math.csv', row.names=FALSE)

```



#Beating the odds
##High poverty schools this year
```{r}
testscores %>% filter(data_year==thisyear, povertycategory=='High') %>% group_by(categoryname) %>% summarise(count=n()) %>% mutate(pct=round(count/sum(count)*100,1))%>%  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```

##high poverty schools in math
```{r}
testscores %>% filter(povertycategory=='High', subject=='M') %>% tabyl(data_year, categoryname) %>% 
  #adorn_totals("col") %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(rounding = "half up", digits = 0) %>%
   adorn_ns() %>% 
   adorn_title("top")%>%
  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```
##high-poverty schools reading
```{r}
testscores %>% filter(povertycategory=="High", subject=='R') %>% tabyl(data_year, categoryname) %>% 
  #adorn_totals("col") %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(rounding = "half up", digits = 0) %>%
   adorn_ns() %>% 
   adorn_title("top")%>%
  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```


##Metro schools - better than expected
```{r message=FALSE}
#exports files with metro high poverty schools

fallingshortmetro <- testscores %>% filter(data_year==thisyear, povertycategory=='High', categoryname=='Falling short', metro7_strib=='YES') %>% select(subject, districtname, schoolname,  pctprof, predicted, residual,  pctpoverty, districttype) %>% arrange((schoolname))

betterthanexpectedmetro <- testscores %>% filter(data_year==thisyear, povertycategory=='High', categoryname=='Better than expected', metro7_strib=='YES') %>%  select(subject, districtname, schoolname,  pctprof, predicted,  pctpoverty, districttype)%>% arrange((schoolname)) %>% mutate(pct_prof= round(pctprof*100,1),                                                                                                  predict = round(predicted*100,1)) %>% select(-pctprof, -predicted) 

#write.csv(fallingshortmetro, "fallingshort_metro.csv", row.names=FALSE)
write.csv(betterthanexpectedmetro, "better_metro.csv", row.names=FALSE)

datatable(betterthanexpectedmetro)
```



##Metro districts -How schools landed

```{r}
testscores %>% filter(subject=='M', metro7_strib=='YES', districttype!='07', data_year==thisyear) %>% tabyl(districtname, categoryname) %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(rounding = "half up", digits = 0) %>%
   adorn_ns() %>% 
   adorn_title("top")%>%
  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```





# Scatterplots
```{r}
#need this for Windows machine to correctly find the Arial font family
windowsFonts(Arial=windowsFont("TT Arial")) 

testscores_math <-  testscores %>% filter(subject=='M', data_year==thisyear, categoryname!='Not enough students tested')

#here's the scatterplot
#notice that we didn't need to make a separate dataframe; we're pulling from the main one
math_scatterplot <- ggplot(testscores_math, aes(x = pctpoverty, y = pctprof)) +
  theme_minimal(base_size = 14, base_family = "Arial") + 
  geom_point(size = 1, alpha = 0.5, aes(color=categoryname)) + 
  scale_y_continuous(labels = percent, limits=c(0,1))+
  scale_x_continuous(labels = percent, limits=c(0,1))+
  geom_smooth(method='lm')+
  #geom_smooth(method = lm, se = FALSE, color = "black", linetype = "dotdash", size = 0.3) +
  labs(title = "Math test scores", 
       caption = "Graphic by MaryJo Webster",
       x="Percent poverty",
       y="Percent proficient")



#plot the scatterplot
plot(math_scatterplot)

scatter_math <-  testscores_math %>% select(schoolid, categoryname, pctpoverty, pctprof) %>% mutate(pctpov = round(pctpoverty*100,2),
                                                                                       prof = round(pctprof*100,2)) %>% select(-pctpoverty, -pctprof)

write.csv(scatter_math, './output/scatter_math.csv', row.names=FALSE)
```


```{r}
windowsFonts(Arial=windowsFont("TT Arial")) 

testscores_read <-  testscores %>% filter(subject=='R', data_year==thisyear, categoryname!="Not enough students tested")

#here's the scatterplot
#notice that we didn't need to make a separate dataframe; we're pulling from the main one
read_scatterplot <- ggplot(testscores_read, aes(x = pctpoverty, y = pctprof)) +
  theme_minimal(base_size = 14, base_family = "Arial") + 
  geom_point(size = 1, alpha = 0.5, aes(color=categoryname)) + 
  scale_y_continuous(labels = percent, limits=c(0,1))+
  scale_x_continuous(labels = percent, limits=c(0,1))+
  geom_smooth(method='lm')+
  #geom_smooth(method = lm, se = FALSE, color = "black", linetype = "dotdash", size = 0.3) +
  labs(title = "Reading test scores", 
       caption = "Graphic by MaryJo Webster",
       x="Percent poverty",
       y="Percent proficient")



#plot the scatterplot
plot(read_scatterplot)



scatter_read <-  testscores_read %>% select(schoolid, categoryname, pctpoverty, pctprof) %>% mutate(pctpov = round(pctpoverty*100,2),
                                                                                       prof = round(pctprof*100,2)) %>% select(-pctpoverty, -pctprof)
write.csv(scatter_read , './output/scatter_read.csv', row.names=FALSE)
```






##How do beating odds schools compare to others?
This shows proficiency for high-poverty beating odds schools<br>
Compare these to the next chart for all high-poverty schools
```{r}
testscores %>% filter(data_year==thisyear, povertycategory=='High', categoryname=='Better than expected') %>% group_by(subject) %>% summarise(prof=sum(numproficient), tot=sum(cnt_tested)) %>% mutate(pct=round((prof/tot)*100,1)) %>% select(subject, pct)%>%  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")


```

##All high-poverty schools
Proficiency percentage
```{r}
testscores %>% filter(data_year==thisyear, povertycategory=='High') %>% group_by(subject) %>% summarise(prof=sum(numproficient), tot=sum(cnt_tested)) %>% mutate(pct=round((prof/tot)*100,1)) %>% select(subject, pct)%>%  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```

#Districts
## math scores overall
```{r}
library(tidyr)
districts_mca <-  districts_mca %>% mutate(pctprof =round((lev3+lev4)/tot_tested*100,2))


dist_math <-  districts_mca %>%  filter(report_category=='All Categories', subject=='M') %>% select(districtid, datayear, pctprof)

dist_math_wide <-  dist_math %>% spread(datayear, pctprof) %>% clean_names()

dist_math_wide <-  left_join(dist_math_wide, district_list %>% select(id_number, organization,district_type, metro7county, location), by=c("districtid"="id_number"))

dist_math_wide <-  dist_math_wide %>% mutate(diff = x18_19-x17_18)

dist_math_wide %>% filter(metro7county=='yes', district_type=='01' | district_type=='03') %>% select(-districtid, -district_type, -metro7county, -location) %>% arrange(desc(diff))%>%  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```

##reading scores
```{r}
dist_read <-  districts_mca %>%  filter(report_category=='All Categories', subject=='R') %>% select(districtid, datayear, pctprof)

dist_read_wide <-  dist_read %>% spread(datayear, pctprof) %>% clean_names()

dist_read_wide <-  left_join(dist_read_wide, district_list %>% select(id_number, organization,district_type, metro7county, location), by=c("districtid"="id_number"))

dist_read_wide <-  dist_read_wide %>% mutate(diff = x18_19-x17_18)

dist_read_wide %>% filter(metro7county=='yes', district_type=='01' | district_type=='03') %>% select(-districtid, -district_type, -metro7county, -location) %>% arrange(desc(diff))%>%  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```



#Duluth
##top level numbers
```{r}
districts_mca %>% filter(districtid=='0709-01-000') %>% select(datayear, subject, report_description, pctprof) %>% arrange(subject, datayear) %>% 
  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```



##Duluth beating the odds
```{r}
testscores %>% filter(districtnumber=='0709', data_year==thisyear) %>% select(data_year, schoolname, subject, povertycategory,categoryname, pctprof ) %>% 
  mutate(pct_prof = round(pctprof*100,1))%>% arrange(desc(pctprof))%>% select(-pctprof) %>%  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```

##Duluth - math
```{r message=FALSE}



duluth_math <- testscores %>% filter(districtnumber=='0709', data_year==thisyear | data_year== lastyear, subject=='M') %>% select(data_year, schoolname,  pctprof ) %>% mutate(pct_prof=round(pctprof*100,1)) %>% select(-pctprof)

duluth_math <-  dcast(duluth_math, schoolname ~ data_year) %>% clean_names() %>% mutate(change=round((x18_19-x17_18),1))

duluth_math %>% arrange(desc(change))%>%  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")

```

##Duluth reading
```{r message=FALSE}
duluth_read <- testscores %>% filter(districtnumber=='0709', data_year==thisyear | data_year== lastyear, subject=='R') %>% select(data_year, schoolname,  pctprof ) %>% mutate(pct_prof=round(pctprof*100,1)) %>% select(-pctprof)

duluth_read <-  dcast(duluth_read, schoolname ~ data_year) %>% clean_names() %>% mutate(change=round((x18_19-x17_18),1))

duluth_read %>% arrange(desc(change))%>%  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```


#Minneapolis schools
##math
```{r message=FALSE}
mps_math <-   testscores %>% filter(districtnumber=='0001', districttype=='03', data_year==thisyear | data_year== lastyear, subject=='M') %>% select(data_year, schoolname,  pctprof ) %>% mutate(pct_prof=round(pctprof*100,1)) %>% select(-pctprof)

mps_math <-  dcast(mps_math, schoolname ~ data_year) %>% clean_names() %>%  mutate(change=round((x18_19-x17_18),1))

datatable(mps_math)


```

## reading
```{r warning=FALSE, message=FALSE}
mps_read <- testscores %>% filter(districtnumber=='0001', districttype=='03', data_year==thisyear | data_year== lastyear, subject=='R') %>% select(data_year, schoolname,  pctprof ) %>% mutate(pct_prof=round(pctprof*100,1)) %>% select(-pctprof)

mps_read <-  dcast(mps_read, schoolname ~ data_year) %>% clean_names() %>% mutate(change=round((x18_19-x17_18),1))

datatable(mps_read)
```



#St Paul schools
## Math
```{r}
sp_math <- testscores %>% filter(districtnumber=='0625', districttype=='01', data_year==thisyear | data_year== lastyear, subject=='M') %>% select(data_year, schoolname,  pctprof ) %>% mutate(pct_prof=round(pctprof*100,1)) %>% select(-pctprof)

sp_math <-  dcast(sp_math, schoolname ~ data_year) %>% clean_names() %>% mutate(change=round((x18_19-x17_18),1))

datatable(sp_math)
```

##Reading
```{r}
sp_r <- testscores %>% filter(districtnumber=='0625', districttype=='01', data_year==thisyear | data_year== lastyear, subject=='R') %>% select(data_year, schoolname,  pctprof ) %>% mutate(pct_prof=round(pctprof*100,1)) %>% select(-pctprof)

sp_r <-  dcast(sp_r, schoolname ~ data_year) %>% clean_names() %>% mutate(change=round((x18_19-x17_18),1))

datatable(sp_r)
```



#Anoka Hennepin
##math
```{r}
ah_math <- testscores %>% filter(districtnumber=='0011', districttype=='01', data_year==thisyear | data_year== lastyear, subject=='M') %>% select(data_year, schoolname,  pctprof ) %>% mutate(pct_prof=round(pctprof*100,1)) %>% select(-pctprof)

ah_math <-  dcast(ah_math, schoolname ~ data_year) %>% clean_names() %>% mutate(change=round((x18_19-x17_18),1))

datatable(ah_math)
```

##reading
```{r}
ah_read <- testscores %>% filter(districtnumber=='0011', districttype=='01', data_year==thisyear | data_year== lastyear, subject=='R') %>% select(data_year, schoolname,  pctprof ) %>% mutate(pct_prof=round(pctprof*100,1)) %>% select(-pctprof)

ah_read <-  dcast(ah_read, schoolname ~ data_year) %>% clean_names() %>%  mutate(change=round((x18_19-x17_18),1))

datatable(ah_read )
```




#Charter schools
includes all charter schools
```{r}

testscores <-  testscores %>% mutate(dis_type_new = case_when(districttype %in% c("07", "70")~'charter',
                                                              districttype %in% c("01", "03")~'trad'))

scores_by_disttype <- testscores %>% group_by(dis_type_new, data_year, subject) %>% 
  summarise(tot_test = sum(cnt_tested),num_lev3 = sum(cntlev3),num_lev4 = sum(cntlev4)) %>% 
  mutate(pct = round(((num_lev3+num_lev4)/tot_test)*100,1)) %>% select(dis_type_new, data_year, subject, pct)




scores_by_disttype <-  dcast(scores_by_disttype, dis_type_new+subject ~ data_year) %>% clean_names() 

scores_by_disttype%>%  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")

```

##By location and type
This only includes elementary, middle/junior and high schools (no ALCs, etc.)

```{r}

scores_by_location<- testscores %>%   group_by(dis_type_new, location_strib, data_year, subject) %>% 
  summarise(tot_test = sum(cnt_tested),num_lev3 = sum(cntlev3),num_lev4 = sum(cntlev4)) %>% 
   mutate(pct = round(((num_lev3+num_lev4)/tot_test)*100,1))  %>% select(location_strib, data_year, subject, pct)




scores_by_location_wide <-  dcast(scores_by_location, dis_type_new+location_strib+subject ~ data_year) %>% clean_names() 

scores_by_location_wide%>%  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```

##By metro and type
Only includes elementaries, middle/junior and high schools (no ALCs, etc.)
```{r}
testscores %>% filter(data_year==thisyear) %>%   group_by(dis_type_new, metro7_strib, subject) %>% 
  summarise(tot_test = sum(cnt_tested),num_lev3 = sum(cntlev3),num_lev4 = sum(cntlev4)) %>% 
   mutate(pct = round(((num_lev3+num_lev4)/tot_test)*100,1))  %>% select(subject, dis_type_new, metro7_strib, pct) %>% arrange(subject, desc(pct))%>%  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```


```{r movetoS3, echo=FALSE, eval=FALSE, results="hide"}




Sys.setenv("AWS_ACCESS_KEY_ID" =  rstudioapi::askForPassword("AWS_ACCESS_KEY_ID"),

           "AWS_SECRET_ACCESS_KEY" = rstudioapi::askForPassword("AWS_SECRET_ACCESS_KEY"))
		   
		   get_bucket("strib-data-public")


put_object(file = "testscore_analysis.html", object = "projects/schools/testscore_analysis.html", bucket = "strib-data-public")

put_object(file = "testscore_analysis.html", object = "projects/schools/testscore_analysis.html", bucket = "strib-data-public", acl=c("public-read"))



```


