---
title: "School test score data"
output: html_notebook
---



```{r}
library(tidyverse)
library(janitor)
library(readxl)
library(RMySQL)
library(lubridate) #date functions
library(scales) #needed for stacked bar chart axis labels
library(knitr) #needed for making tables in markdown page
library(aws.s3)
library(htmltools)#this is needed for Rstudio to display kable and other html code
library(rmarkdown)
library(DT) #needed for making  searchable sortable data tble
library(kableExtra)
library(ggthemes)
library(stringr)
library(formattable)
library(broom)

# use R script to import new data from the Excel file the state produces
# that spits out a csv file
# import that csv file to the mca2019_present table in mySQL server



#set this path for the new PUBLIC enrollment file
public_enroll_file <-  './data/enrollment_public_file_2023.xlsx'


#set this to the most current year of data
dtyr='22-23'

#set this to the previous year of data
prevdtyr = '21-22'
```







#import data from mySQL  (be sure to add the new data first)
```{r}
con <- dbConnect(RMySQL::MySQL(), host = Sys.getenv("host"), dbname="Schools",user= Sys.getenv("userid"), password=Sys.getenv("pwd"))

data_a <- dbSendQuery(con, "select * from DistrictList")
district_list<- fetch(data_a, n=-1)
district_list <-  district_list %>% clean_names()
dbClearResult((data_a))



#Pull SchoolList (needs to use distinct because there are some duplicates)
data_b <- dbSendQuery(con, "select distinct schoolid, district_type, schoolnumber, district_name, school_name, 
                     school_classification, school_type, grades, metro7_strib, location_strib, city 
                     from schoollist")
school_list <- fetch(data_b, n=-1)
school_list <-  school_list %>% clean_names()

dbClearResult(data_b)

data_c <-  dbSendQuery(con, "select data_year, county_number, county_name, district_number, district_type, district_name, school_number, school_name, school_classification, test_name, grade, subject, group_category, student_group, total_tested, filter_all, count_level_d, count_level_p, count_level_m, count_level_e, count_absent, count_invalid, count_invalid_student_behavior, count_invalid_device, count_invalid_other, count_medical_exempt, count_not_attempted, count_not_complete, count_not_enrolled, count_refused, count_refused_parent, count_refused_student, count_wrong_grade, count_valid_scores_mca_with_accommodations, count_extenuating_circumstances_attempted, count_extenuating_circumstances_not_attempted, level, idnumber  
                       from mca2019_present")


mca_mtas_original <-  fetch(data_c, n= -1)

dbClearResult(data_c)


# POVERTY DATA; this assumes the most current year of poverty data is in the SQL table "enroll"
data_d <- dbSendQuery(con, "select idnumber, data_year, total_students_eligible_for_free_or_reduced_priced_meals_count, total_students_eligible_for_free_or_reduced_priced_meals_percent, total_enrollment
from enroll where upper(grade)='ALL GRADES' and endyr>=2018")

poverty <- fetch(data_d, n= -1)

dbClearResult(data_d)



#disconnect connection
dbDisconnect(con)


school_list <-  school_list %>% rename(idnumber=schoolid, metro7county=metro7_strib, location = location_strib, standard_school_name=school_name)

district_list <-  district_list %>% rename(idnumber=id_number, standard_district_name=organization)


#create a single file with both schools and districts to join to the mca_mtas data
master_list <-  bind_rows(school_list %>% select(idnumber, district_type, standard_district_name=district_name, standard_school_name, metro7county, location, school_classification, school_type, grades, city), district_list %>% select(idnumber, district_type, standard_district_name, metro7county, location))






  
```


```{r}

#before running next code, check how many records are in mca_mtas_original; then run this code and make sure you get the same number of records

#join school and district info to mca_mtas
mca_mtas_original <-  left_join(mca_mtas_original, master_list %>% select(-district_type, -school_classification), by=c("idnumber"="idnumber"))


```




# clean up poverty data
```{r}
# remove percentage signs (they appear randomly in the file)
poverty <- poverty %>%  mutate(pct_poverty = str_remove(total_students_eligible_for_free_or_reduced_priced_meals_percent, "%"))


public_poverty   <- 
  poverty %>% 
  select(data_year, idnumber ,total_enrollment, poverty = total_students_eligible_for_free_or_reduced_priced_meals_count, pct_poverty) %>% 
  mutate(poverty_calc = case_when(pct_poverty=='>=90' ~ .90,   #this creates a numeric field for use in the regression
                                  pct_poverty=='0.00' ~ 0,
    pct_poverty!='NA' ~ poverty/total_enrollment),
   
         povertylevel = case_when(pct_poverty=='>=90' ~ 'high',   #this creates a category for use in analysis
                                  poverty_calc >=.6 ~ 'high',
                                  poverty_calc<.6 & poverty_calc>.3 ~ 'medium',
                                  poverty_calc<=.3  ~ 'low',
                                  TRUE ~ 'unk'),
    pct_pov_public = case_when(pct_poverty=='NA'~'Not public',  #this creates a text field for display online
                               pct_poverty=='>=90' ~ '>=90%',
                                
                                TRUE~pct_poverty)) 




```



```{r}
#join poverty info to mca_mtas; this new table should have same number of records as the mca_mtas_original
mca_mtas <-  left_join(mca_mtas_original, public_poverty, by=c("idnumber"="idnumber", "data_year"="data_year"))




#clean leading/trailing spaces on key fields
mca_mtas <-  mca_mtas %>%
  mutate(grade= str_trim(grade), 
         subject=str_trim(subject), 
         student_group=str_trim(student_group))



#fill out the numproficient column
mca_mtas <-  mca_mtas %>% mutate(numproficient = count_level_m+count_level_e,
                                 pct_proficient = numproficient/total_tested)


#add uniqueID, adjust datayear and grades fields
mca_mtas <- mca_mtas %>% mutate(uniqueID=paste(idnumber,'-',substr(data_year,1,2),'to',substr(data_year,4,6),'-',subject), 
                                    datayear_new=paste(substr(data_year,1,2),'to',substr(data_year,4,6)), 
                                    grades = str_trim(grades),
                                    grades_new= case_when(str_length(grades)==1 ~ grades, TRUE~str_replace(grades, '-', ' to ')))


# add a school_type column that converts the classification to words
mca_mtas <- mca_mtas%>% mutate(school_type= case_when(school_classification=='10'~'Elementary (PK-6)',
                                                         school_classification=='20'~ 'Middle School (5-8)',
                                                         school_classification=='31'~'Junior High (7-8 or 7-9)',
                                                          school_classification=='33'~'Secondary (7-12)',
                                                          school_classification=='32'~'Senior High (9-12)',
                                                         school_classification=='40'~'Elem/Sec Combo (K-12)',
                                                         school_classification=='46'~'Distance learning'))





```


# Schools where we don't have poverty rates for current year (in some cases, we do for earlier years)
```{r}
mca_mtas %>% filter(pct_pov_public=='Not public', level=='school', grade=='00', student_group=='All students', data_year=='22-23', filter_all=='N', total_tested>=25)
```


# Rochester didn't give tests in 2023 due to technical problems
```{r}
mca_mtas %>% filter(district_number=='0535', level=='school', grade=='00', student_group=='All students', filter_all=='N') 
```


```{r}

# this pulls out Rochester data in order to add it back in as part of the regression analysis
# Rochester didn't test in 2023
#won't need this in future years
#there are a few other spots where rochester is added in that needs to be edited in future years

rochester <- mca_mtas %>% filter(district_number=='0535',level=='school', filter_all=='N', student_group=='All students', grade=='00')%>% mutate(notes='Tests not given in 2023 due to technical problems')
```




```{r}


#this section is just for the lookup tool that goes with the first day story (not the regression analysis)



#NEED TO WINNOW DOWN TO JUST SCHOOLS THAT EXISTED IN CURRENT YEAR


mca_mtas_current_yr <-  mca_mtas %>%   filter(data_year==dtyr) %>% group_by(idnumber) %>% summarize(count=n())

mca_mtas_public <- inner_join(mca_mtas, mca_mtas_current_yr %>% select(idnumber), by=c("idnumber"="idnumber"))


#check to see that we have the right number of schools in the mca_mtas_public table
#this query should yield the same number of records as in the mca_mtas_current_yr table
mca_mtas_public %>% group_by(idnumber) %>% summarise(count=n())





# file for day 1 website lookup

mca_online <-  mca_mtas_public %>% 
  filter(level=='school', data_year %in% c(dtyr, prevdtyr), school_classification!='41', school_classification!='42', school_classification!='43', student_group=='All students') %>% 
  select(uniqueID, schoolid=idnumber, districtnumber=district_number, districttype=district_type, school_number, districtname=standard_district_name, schoolname=standard_school_name, schoolclassification=school_classification, school_type,
         grades_new, metro7_strib=metro7county, location_strib=location, datayear_new, subject, filter_all, grade, cnt_tested=total_tested, cntlev1=count_level_d, cntlev2=count_level_p, cntlev3=count_level_m,
         cntlev4=count_level_e, numproficient, pctprof=pct_proficient, total_enrollment, pct_pov_private=poverty_calc, pct_pov_public,     
         povertycategory=povertylevel, city, level) %>% 
  mutate(notes = case_when(filter_all=='Y' ~ 'Data suppressed; less than 10 students tested',
                           TRUE ~ ''))
  
#write.csv(mca_online, './output/mca_online.csv', row.names=FALSE)


#check the file output to make sure it looks ok
mca_online %>% filter(schoolid=='0001-03-105') %>% select(datayear_new, subject, grade, pctprof, city) %>% arrange(datayear_new, subject, grade)

```




# Clean up the opt out fields
```{r}

mca_mtas$count_invalid[is.na(mca_mtas$count_invalid)] <- 0
mca_mtas$count_invalid_student_behavior[is.na(mca_mtas$count_invalid_student_behavior)] <- 0
mca_mtas$count_invalid_other[is.na(mca_mtas$count_invalid_other)] <- 0
mca_mtas$count_medical_exempt[is.na(mca_mtas$count_medical_exempt)] <- 0
mca_mtas$count_not_attempted[is.na(mca_mtas$count_not_attempted)] <- 0
mca_mtas$count_not_complete[is.na(mca_mtas$count_not_complete)] <- 0
mca_mtas$count_not_enrolled[is.na(mca_mtas$count_not_enrolled)] <- 0
mca_mtas$count_refused[is.na(mca_mtas$count_refused)] <- 0
mca_mtas$count_refused_parent[is.na(mca_mtas$count_refused_parent)] <- 0
mca_mtas$count_refused_student[is.na(mca_mtas$count_refused_student)] <- 0
mca_mtas$count_wrong_grade[is.na(mca_mtas$count_wrong_grade)] <- 0
mca_mtas$count_extenuating_circumstances_attempted[is.na(mca_mtas$count_extenuating_circumstances_attempted)] <- 0
mca_mtas$count_extenuating_circumstances_not_attempted[is.na(mca_mtas$count_extenuating_circumstances_not_attempted)] <- 0
#not including count_invalid and count_refused as those fields are no longer used
#not including count_wrong_grade and count_not_enrolled since those are kids who weren't supposed to be in this group to begin with


# add a total_students column 
#and columns counting kids not taking tests due to covid or other circumstances
mca_mtas <-  mca_mtas %>% mutate(total_students = total_tested + count_absent +  count_invalid_student_behavior+
                                   count_invalid_device+count_not_attempted+count_medical_exempt + count_not_complete+
                                    count_refused_parent+count_refused_student + count_extenuating_circumstances_attempted + count_extenuating_circumstances_not_attempted,
                                 
                                 not_complete_non_covid = count_absent +  count_invalid_student_behavior+
                                   count_invalid_device+count_not_attempted+count_medical_exempt + count_not_complete+count_refused_parent+count_refused_student,
                                 covid =count_extenuating_circumstances_attempted + count_extenuating_circumstances_not_attempted )



# export a state level file with results for each student group, plus opt out rates
state_level_all_years <- mca_mtas %>% filter(level=='state') %>% select(data_year, grade, subject, group_category, student_group, total_tested, count_level_d, count_level_p, count_level_m, count_level_e, numproficient, pct_proficient, count_absent, count_invalid, count_invalid_student_behavior, count_invalid_device, count_invalid_other, count_medical_exempt, count_not_attempted, count_not_complete, count_wrong_grade, count_valid_scores_mca_with_accommodations, count_extenuating_circumstances_attempted, count_extenuating_circumstances_not_attempted, total_students, not_complete_non_covid, covid) %>% 
  mutate(opt_out_percentage = (not_complete_non_covid+covid)/total_students)

write.csv(state_level_all_years, './output/state_level_all_years.csv', row.names=FALSE)
```




# regression analysis

```{r}
#math results, all students, all grades collapsed, school level, 25 or more students tested, not filtered, and where we have poverty calculation

math <- mca_mtas %>% filter(subject=='MATH', grade=='00', level=='school', student_group=='All students', filter_all=='N', total_tested>=25, poverty_calc!='NA') 


#reading results with same filters
read <- mca_mtas %>% filter(subject=='Reading', grade=='00', level=='school', student_group=='All students', filter_all=='N', total_tested>=25, poverty_calc!='NA')


#export files for regression analysis
#write.csv(math, './data/math_regression.csv', row.names=FALSE)
#write.csv(read, './data/read_regression.csv', row.names = FALSE)


#create two files of the schools excluded from the analysis because total_tested<25 or no poverty calculation
#these need to be added back to the final data file


#excluded due to less than 25 tested
excluded1 <-  mca_mtas %>% filter( grade=='00', level=='school', student_group=='All students', filter_all=='N',total_tested<25 )%>%
  mutate(notes='Less than 25 students tested', predicted=NA_real_,  residual=NA_real_)


excluded2 <- mca_mtas %>% filter(grade=='00', level=='school', student_group=='All students', filter_all=='N', is.na(poverty_calc)) %>%
  mutate(notes='Poverty data not available', predicted=NA_real_,  residual=NA_real_)


#schools that didn't have test result data due to suppression
missing <- mca_mtas %>% filter(level=='school', filter_all=='Y', student_group=='All students', grade=='00')%>% mutate(notes='Less than 10 students tested')





#combine all the files  (remove Rochester in future years)
combined <-  bind_rows(math, read, excluded1, excluded2, missing, rochester)



#test to make sure all the schools are here; this query should result in zero

anti_join(mca_mtas %>% filter(level=='school', grade=='00', student_group=='All students'), combined, by=c("idnumber"="idnumber", "data_year"="data_year", "subject"="subject"))



```




```{r}

# MATH REGRESSION ---------------------------------------------------------

#build model
math_model <- lm(pct_proficient ~ poverty_calc, data=math)

#predicted scores
pred_math <- predict(math_model, math)



#add predicted value
math <-  math %>%  mutate(predicted=pred_math)



#add residual
math <-  math  %>% mutate(residual = pct_proficient-predicted)

#summary(math_model)



# READING REGRESSION ------------------------------------------------------

#build model
read_model <- lm(pct_proficient ~ poverty_calc, data=read)

#predicted scores
pred_read <- predict(read_model, read)

#add predicted value
read <-  read %>%  mutate(predicted=pred_read)

#add residual
read <-  read  %>% mutate(residual = pct_proficient-predicted)

```

```{r}
augment(math_model, data = math) %>% 
  arrange(.fitted)


#this calculates the fit line needed for scatterplots
#math fit line
fit <-  augment(math_model, data = math) %>% group_by(.fitted) %>% summarise(count=n())

write.csv(fit, 'math_fit_line.csv', row.names = FALSE)



#reading fit line
fit_read <-  augment(read_model, data = read) %>% group_by(.fitted) %>% summarise(count=n())

write.csv(fit_read, 'read_fit_line.csv', row.names = FALSE)
```


```{r}


augment(math_model, data=math) %>% 
  ggplot() +
  geom_point(aes(x = poverty_calc, y = pct_proficient), size = .8) + 
  geom_line(aes(x = poverty_calc, y = .fitted), color = "blue") + #the fitted line
  #geom_point(data=math2223 %>% filter(idnumber=='0837-01-020'), aes(x = poverty_calc, y = pct_proficient), size = 2, color="red")
  labs(x = "Pct Poverty", y = "Proficiency") +
  theme_minimal()
```





```{r}
augment(read_model, data=read) %>% 
  ggplot() +
  geom_point(aes(x = poverty_calc, y = pct_proficient), size = .8) + 
  geom_line(aes(x = poverty_calc, y = .fitted), color = "blue") + #the fitted line
  geom_point(data=read %>% filter(idnumber=='0837-01-020'), aes(x = poverty_calc, y = pct_proficient), size = 2, color="red")
  labs(x = "Pct Poverty", y = "Proficiency") +
  theme_minimal()
```


```{r}

# add categories based on the regression results
math <- math %>% mutate(categorynum= case_when(residual==NA_real_ ~0, 
                                                           residual<  -0.0951 ~1,
                                                           between(residual, -.0951, .09509)~2,
                                                           residual>  0.09509 ~3, 
                                                           TRUE ~99))

math <-  math %>% 
  mutate(categoryname= case_when(categorynum==99~"Not enough students tested", 
                                 categorynum==1~"Falling short", 
                                 categorynum==2~"As expected",
                                 categorynum==3~"Better than expected", 
                                 TRUE~"99"))



read <- read %>% mutate(categorynum= case_when(residual==NA_real_ ~0, 
                                                           residual<  -0.0951 ~1,
                                                           between(residual, -.0951, .09509)~2,
                                                           residual>  0.09509 ~3, 
                                                           TRUE ~99))

read <-  read %>% 
  mutate(categoryname= case_when(categorynum==99~"Not enough students tested", 
                                 categorynum==1~"Falling short", 
                                 categorynum==2~"As expected",
                                 categorynum==3~"Better than expected", 
                                 TRUE~"99"))









#union math, read, math_excluded, read_excluded -- call new table testscores\ (remove Rochester in future years)

testscores <- bind_rows(read, math, excluded1, excluded2, missing, rochester) 






# create file for Beating the odds online tool
testscores_public <- testscores %>% 
  select(uniqueID, schoolid=idnumber, districtnumber=district_number, districttype=district_type, school_number, districtname=standard_district_name, schoolname=standard_school_name, schoolclassification=school_classification, school_type,
         grades_new, metro7_strib=metro7county, location_strib=location, datayear_new, subject, cnt_tested=total_tested, cntlev1=count_level_d, cntlev2=count_level_p, cntlev3=count_level_m,
         cntlev4=count_level_e, numproficient, pctprof=pct_proficient, total_enrollment, 
         pct_pov_private=poverty_calc, 
         pct_pov_public, 
         predicted, residual, notes, categorynum, 
         categoryname, povertycategory=povertylevel)

#set these aside to add back in later; DON'T NEED THIS IN FUTURE YEARS; Rochester didn't test in 2023 due to technical problems
rochester2 <-  testscores_public %>% filter(districtnumber=='0535')



#NEED TO WINNOW DOWN TO JUST SCHOOLS THAT EXISTED IN CURRENT YEAR
#also filter to elementary schools, secondary schools and online schools

testscores_current_yr <-  testscores %>%   filter(data_year==dtyr) %>% group_by(idnumber) %>% summarize(count=n())

testscores_public <- inner_join(testscores_public, testscores_current_yr %>% select(idnumber), by=c("schoolid"="idnumber")) 


#need to add Rochester back in because they didn't test students in 2023 (exclude this in future years; but keep the school classification part)
testscores_public <-  bind_rows(testscores_public, rochester2)%>% 
  filter(schoolclassification %in% c( '10', '20', '32', '33', '46', '31', '40'))




#export CSV for data visualization
write.csv(testscores_public, "./output/beating_odds_output.csv", row.names = FALSE)



```





# Opt outs

```{r}

mca_mtas %>% filter(level=='state', student_group=='All students', grade=='00') %>% group_by(data_year) %>% 
  summarise( tottested = sum(total_tested), totstudents = sum(total_students), non_covid =sum(not_complete_non_covid), cov = sum(covid)) %>% 
  mutate(pct_not_tested = (non_covid+cov)/totstudents) %>% 
  arrange(desc(pct_not_tested))


```


# Opt outs grade 11 math

```{r}

mca_mtas %>% filter(level=='state', student_group=='All students', grade=='11') %>% group_by(data_year) %>% 
  summarise( tottested = sum(total_tested), totstudents = sum(total_students), non_covid =sum(not_complete_non_covid), cov = sum(covid)) %>% 
  mutate(pct_not_tested = (non_covid+cov)/totstudents) %>% 
  arrange(desc(pct_not_tested))

```



# opt outs by student group - math
```{r}
mca_mtas %>% filter(level=='state', subject=='MATH', data_year==dtyr) %>% group_by(student_group) %>% 
  summarise( tottested = sum(total_tested), totstudents = sum(total_students), non_covid =sum(not_complete_non_covid), cov = sum(covid)) %>% 
  mutate(pct_not_tested = (non_covid+cov)/totstudents) %>% 
  select(student_group, pct_not_tested) %>% 
  arrange(desc(pct_not_tested))
```



# Statewide achievement

```{r, warning=FALSE, message=FALSE}
mca_mtas %>%
  filter( student_group=='All students',  level=='state', grade=='00') %>% 
mutate(pctprof = percent((count_level_m+count_level_e)/total_tested)) %>% 
  select(data_year, subject, pctprof) %>% 
  arrange(subject, data_year) %>% 
  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")


```

# proficiency by grade - reading
```{r}
mca_mtas %>%
  filter( student_group=='All students',  level=='state', subject=='Reading') %>% 
mutate(pctprof = percent((count_level_m+count_level_e)/total_tested)) %>% 
  select(data_year, grade, pctprof) %>% 
  arrange(grade, data_year) %>% 
  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")


write.csv(mca_mtas %>%
  filter( student_group=='All students',  level=='state', subject=='Reading') %>% 
mutate(pctprof = percent((count_level_m+count_level_e)/total_tested)) %>% 
  select(data_year, grade, pctprof), './output/Reading_by_grade_all_years.csv', row.names=FALSE)
```


# proficiency by grade -- math
```{r}
mca_mtas %>%
  filter( student_group=='All students',  level=='state', subject=='MATH') %>% 
mutate(pctprof = percent((count_level_m+count_level_e)/total_tested)) %>% 
  select(data_year, grade, pctprof) %>% 
  arrange(grade, data_year) %>% 
  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")



write.csv(mca_mtas %>%
  filter( student_group=='All students',  level=='state', subject=='MATH') %>% 
mutate(pctprof = percent((count_level_m+count_level_e)/total_tested)) %>% 
  select(data_year, grade, pctprof), './output/MATH_by_grade_all_years.csv', row.names=FALSE)
```




## Race groups - math
```{r, warning=FALSE, message=FALSE}

state_race_math <-  mca_mtas %>%
  filter(group_category %in% c('Race/Ethnicity', 'Federal Race/Ethnicity', 'State Race/Ethnicity'), level=='state', toupper(subject)=='MATH', grade=='00') %>% 
  select(student_group, data_year, count_level_m, count_level_e, total_tested) %>% 
  mutate(pctprof = (count_level_m+count_level_e)/total_tested) %>% 
  arrange(student_group, data_year)



g_race_math <- ggplot(data=state_race_math, aes(x = state_race_math$data_year, y = pctprof, group=student_group, col=student_group)) +
                  geom_line(stat="identity",  size=1)+
  scale_y_continuous(name="Percent proficient", limits=c(0, .8), labels=percent)+
  theme(legend.key.size = unit(.3, "cm"), legend.position="top", legend.title = element_blank())+
  scale_x_discrete(name="School year")+
    theme_hc()+
      labs(title = "Achievement in math")
plot(g_race_math)
```




```{r}

state_race_read <-  mca_mtas %>%
  filter(group_category %in% c('Race/Ethnicity', 'Federal Race/Ethnicity', 'State Race/Ethnicity'), level=='state', toupper(subject)=='READING', grade=='00') %>% 
  select(student_group, data_year, count_level_m, count_level_e, total_tested) %>% 
  mutate(pctprof = (count_level_m+count_level_e)/total_tested) %>% 
  arrange(student_group, data_year)


g_race_read <- ggplot(data=state_race_read, aes(x = data_year, y = pctprof, group=student_group, col=student_group)) +
                  geom_line(stat="identity",  size=1)+
  scale_y_continuous(name="Percent proficient", limits=c(0, .8), labels=percent)+
  theme(legend.key.size = unit(.3, "cm"), legend.position="top", legend.title = element_blank())+
  scale_x_discrete(name="School year")+
    theme_hc()+
      labs(title = "Achievement in reading")
plot(g_race_read)
```


```{r}
subgroups<-  mca_mtas %>%
  filter(level=='state', grade=='00', filter_all=='N') %>% 
  select(subject, group_category, student_group, data_year,  count_level_m, count_level_e, total_tested) %>% 
  mutate(pctprof = (count_level_m+count_level_e)/total_tested)

write.csv(subgroups, './output/subgroups.csv', row.names = FALSE)

subgroups %>% select(group_category, student_group, data_year, subject, pctprof) %>% arrange(subject, group_category, student_group, data_year )
```


# Gender- reading
```{r}
gender <-  mca_mtas %>% filter(group_category=='Gender', grade=='00', subject=='Reading', level=='state') %>% select(data_year, student_group, count_level_m, count_level_e, total_tested) %>% 
  mutate(pctprof = (count_level_m+count_level_e)/total_tested)

ggplot(data=gender, aes(x=data_year, y=pctprof, group=student_group, col=student_group))+
    geom_line(stat="identity")+
  scale_y_continuous(name="Percent proficient", limits=c(0, .8), labels=percent)


mca_mtas %>% group_by(subject) %>% summarise(count=n())
```


# District types
```{r}
mca_mtas %>%   
filter( student_group=='All students',  level=='district', subject=='Reading', filter_all=='N') %>% 
  group_by(data_year, district_type) %>% 
  summarise(levm = sum(count_level_m), leve= sum(count_level_e), tot = sum(total_tested)) %>% 
mutate(pctprof = percent((levm+leve)/tot)) %>% 
  arrange(district_type, data_year) %>% 
  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")

```



# Virtual schools
```{r}
mca_mtas %>%   
filter( student_group=='All students',  level=='school',  filter_all=='N', school_classification=='46') %>% 
  group_by(data_year, subject) %>% 
  summarise(levm = sum(count_level_m), leve= sum(count_level_e), tot = sum(total_tested)) %>% 
mutate(pctprof = percent((levm+leve)/tot)) %>% 
  arrange(subject,  data_year) %>% 
  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```





# find the 10 largest metro districts
```{r}
mca_mtas %>% filter(data_year=='22-23', subject=='MATH', level=='district', metro7county=='yes', grade=='00', student_group=='All students', district_type %in% c("01", "03")) %>% 
  select(idnumber, standard_district_name, numproficient, total_tested, pct_proficient, total_students) %>%
  arrange(desc(total_students))
```


#export data for previous two years for those 10 districts, both math and reading -- for a print table
```{r}
top10metro <-  
mca_mtas %>% 

filter(idnumber %in% c("0011-01-000","0196-01-000", "0625-01-000", "0001-03-000", "0279-01-000", "0833-01-000", "0284-01-000", "0276-01-000", 
                       "0194-01-000", "0621-01-000"),
       data_year %in% c("18-19"),
       level=='district', grade=='00', student_group=='All students') %>% 
  select(idnumber, data_year, subject, standard_district_name, numproficient, pct_proficient, total_tested) %>% 
  arrange(idnumber, subject, data_year)

write.csv(top10metro, './output/top10metro_print_table_18_19.csv', row.names=FALSE)
```

